---
title: "Mini Project 6"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
pro_canc <- read.csv(file = "prostate_cancer.csv", header=TRUE, sep=",")
attach(pro_canc)
```

When seeing the boxplot below for the dataset, the variance is high. So we are performing log transformation.

```{r}
boxplot(pro_canc$psa)
```
```{r}
y <- log(pro_canc$psa, base = exp(1))
boxplot(y)
```

Lets try to 

```{r}
plot(pro_canc$cancervol, y)
fit1 <- lm(y ~ cancervol, data = pro_canc)
abline(fit1)
```
```{r}
plot(pro_canc$weight, y)
fit2 <- lm(y ~ weight, data = pro_canc)
abline(fit2)
```
```{r}
plot(pro_canc$age, y)
fit3 <- lm(y ~ age, data = pro_canc)
abline(fit3)
```
```{r}
plot(pro_canc$benpros, y)
fit4 <- lm(y ~ benpros, data = pro_canc)
abline(fit4)
```
```{r}
plot(pro_canc$vesinv, y)
fit5 <- lm(y ~ factor(vesinv), data = pro_canc)
abline(fit5)
```
```{r}
plot(pro_canc$capspen, y)
fit6 <- lm(y ~ capspen, data = pro_canc)
abline(fit6)
```
```{r}
plot(pro_canc$gleason, y)
fit7 <- lm(y ~ gleason, data = pro_canc)
abline(fit7)
```
```{r}
cor(y, pro_canc$cancervol)
cor(y, pro_canc$vesinv)
cor(y, pro_canc$gleason)
cor(y, pro_canc$capspen)
cor(y, pro_canc$age)
cor(y, pro_canc$benpros)
cor(y, pro_canc$weight)
```

From the plots of individual predictors , the predictors are ordered in decreasing order of their significance with their correlation values listed above

Now ANOVA analysis is performed by adding predictors one by one
```{r}
fit8 <- lm(y ~ cancervol+ factor(vesinv), data = pro_canc)
anova(fit1,fit8)
```
```{r}
summary(fit1)
```
```{r}
summary(fit8)
```

From the aboves summary data of fit 1 and fit 8, we can see that adjusted R-squared for fit8 > fit1. So we keep fit8 for further analysis. Now we continue the ANOVA analysis for next predictor
```{r}
fit9 <- lm(y ~ cancervol+ factor(vesinv)+ gleason, data = pro_canc)
anova(fit8,fit9)
```

We can see that the p-value is less than 0.05 on adding gleason to model, so gleason is a significant predictor.

```{r}
summary(fit9)
```

Also, we can see that Adjusted R-squared for fit9 > fit8, so we can retain fit 9 for further analysis.

```{r}
fit10 <- lm(y ~ cancervol+ factor(vesinv) + gleason + capspen, data = pro_canc)
anova(fit9,fit10)
```

We can see that p-value > 0.05 so capspen is not significant to add in the model. Therefore fit9 is retained for further analysis.

```{r}
fit11 <- lm(y ~ cancervol+ factor(vesinv)  + gleason + age, data = pro_canc)
anova(fit9,fit11)
```

We can see that p-value > 0.05 for age as well. So age is not significant to add in the model. We will keep fit9 and proceed with other predictors.

```{r}
fit12 <- lm(y ~ cancervol+ factor(vesinv)  + gleason + benpros, data = pro_canc)
anova(fit9,fit12)
```

We can see that P- value is less than 0.05 so benpros is significant in the model.

```{r}
summary(fit12)
```

But the value of Adjusted R-squared for fit12 > fit9. So we will continue analysis with fit12.

```{r}
fit13 <- lm(y ~ cancervol+ factor(vesinv)  + gleason + benpros + weight, data = pro_canc)
anova(fit12,fit13)
```

We can see that p-value > 0.05 thus weight is not significant to add in the model. 
From the above analysis we can see that fit12 is the best model.

Now we will verify the above result by comparing fit12=lm(y ~ cancervol+ factor(vesinv)  + gleason + benpros, data = pro_canc) with automatic stepwise model selection procedures 
based on AIC

```{r}
fit12.forward <- step(lm(y ~ 1, data = pro_canc),scope = list(upper = ~cancervol + factor(vesinv) + gleason + benpros + weight + age + capspen),direction ="forward")
```

From above, we can see that y ~ cancervol + gleason + benpros + factor(vesinv) is the best model from AIC feed forward. This confirms the results obtained by the manual computation of best model.

Model assumptions verification:

Residual plot for fit 12 -

```{r}
plot(fitted(fit12), resid(fit12))
abline(h=0)
```

The residual plot shows constant variability.

QQ-Plot for fit12 -

```{r}
qqnorm(resid(fit12))
qqline(resid(fit12))
```

The QQ-Plot verifies the normality assumption.

Prediction of PSA level of patients.

The best model identified is fit12 <- lm(y ~ cancervol+ factor(vesinv)  + gleason + benpros, data = pro_canc)

Here cancervol, gleason and benpros are quantitative variables whereas vesinv is a categorical attribute.

The most frequent category of vesinv is calculated as below.

```{r}
table(vesinv)
```
0 is the most frequent category.

For the other attributes we use mean to calculate predicted value.

```{r}
summary(fit12)
```

The prediction using fit12 model is done as below.

```{r}
exp(-0.65013 + 0.06488*mean(cancervol) + 0.68*0 + 0.33376*mean(gleason) + 0.09136*mean(benpros))
```

